---
title: "Starling Detector Notes"
author: "Vlad Obo"
date: "10/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

source("birdsLib.R")

```

## About this document

## Problem Description

* We have massive sound records  and need to find if there are starling calls

* There are a data set of starling calls 1000+ records

* We think the data in hands reasonably cover the real conditions

* we develop specific application for starlings

* we consider mel frequency cepstrum coeffisients as a feature and HTK and KERAS as ML tools


## Parts

### Choosing MFCC parameters

After playing the data the following numbers were chosen as MFCC's parameters. 

Happy to explaine the choice.

```{r, echo=TRUE}
hopt

wint 

getMel
```



### Bird call detector

* Simple detector to cut birds calls from lengthy sound records would be good to have

* I called it "MFCC power"

* The function is as below

```{r}
getMCPow
```

* How it works. Using 15 seconds from joined starling decreasin calls as an example

```{r}
sigDir="../"
fn="DecreasingCallLargeTmp.wav"
w<-readWave(paste0(sigDir, fn), from = 0, to = 15, units = "sec")

plot(w)

viewSpec(w)

```

```{r}
ss<-getMel(w)

tt<-(1:nrow(ss)) * hopt
matplot(tt, ss, type="l", main="MFCCs, all")
```



```{r}
pov=getMCPow(ss)
plot(pov, type="b", main="MFFC Power")

showMCPow(pov)
title(main="Median filter")
```


```{r}
win.range<-c(0.4, 0.8)
```


* I use this file `r fn` to produce the estimate for target window length

* The target window size set to  `r win.range` sec

```{r}
wins<-readWave(paste0(sigDir, fn)) %>%
  getMel %>%
  getMCPow %>%
  getAWins

ll<-with(wins, end-start)

plot(sort(ll))
abline(h=win.range, lty=2, col="red")


```




## Real audio


```{r, echo=FALSE}
sigDir<-"/Volumes/VladBackup/Birds/StarlingCalls/Ron Sinclair_Turretfield_2009/"
lfn<-"MU000_20091021_172000.wav"
ifn=paste0(sigDir, lfn)

sta=0
sto=30
wL<-mono(readWave(ifn, from=sta, to=sto, units="minutes"))

```

* Now worrking on the file  `r lfn` (the first helf hour only here to avoid memory problems)

* From `r sta` to `r sto`, in minutes

```{r, echo=FALSE}
plot(wL, main=lfn, xlab="Time, s")

pov <- getMel(wL) %>%  getMCPow 
  
pov %>% showMCPow(tit=lfn)

showMCPow(pov, xl = c(1, 100), tit="first 100s")
showMCPow(pov, xl = c(40, 80))
showMCPow(pov, xl = c(40, 60))
plot(wL, xlim=c(40,60))
```

## Action windows in real audio file

* We seek time windows with enough action in them

* It means when the median filter output is greater than the threshold

* Threshold used by default (see function "getAWins") is the top quartile of "pov"

```{r}
## be careful -- remove the tmp file when changing the input wav
tmp.fn="action.wins.RData"
#file.remove(tmp.fn)
if (file.exists(tmp.fn)) load(tmp.fn) else {
  
action.wins <- (ppov<- paste0(sigDir, lfn) %>%
                  readWave(from=0, to = 30, units="min") %>%
                  mono %>%
                  getMel() %>%
                  getMCPow()
                ) %>%
  getAWins()
save(ppov, action.wins, file=tmp.fn) } 

ll<-with(action.wins, end-start)

plot(sort(ll), log="y")
abline(h=win.range, col="red", lty=2)


ii<-which(inrange(ll, win.range))
```


* Number of action windows similar to starling decr.call window size
```{r}
table(inrange(ll, win.range))
```


* Total sum of lengths for the windows of these sizes in sec (this is from the record of 1h)

```{r}
sum(ll[ii])
```

* Let's generate wav including only fragments determined by the action windows. Accomponying selection table is written too. The can be viewed by Raven

```{r}
ofn="wins.wav"
winsWav(infile=ifn, outfile=ofn, ww=action.wins[ii,])

```


* Selection table to be used with original file to show these selected windows

```{r}

writeSelectionTable(wins=action.wins, ofn="actionWins.txt")

```






<!-- ## Test file -->

<!-- ```{r} -->
<!-- sigDir<-"/Volumes/VladBackup/Birds/StarlingCalls/" -->
<!-- lfn<-"ma5ho-ehcz5.L.wav" -->
<!-- wL<-mono(readWave(paste0(sigDir, lfn))) -->

<!-- plot(wL, main=lfn) -->
<!-- viewSpec(wL, main=lfn) -->

<!-- ss=getMel(wL) -->
<!-- pov=getMCPow(ss) -->
<!-- showMCPow(pov, tit=lfn) -->
<!-- ``` -->


* Make